---
title: "Analysis"
author: "Susie and Teddie"
format: pdf
---
Packages: 

```{r}
library(tidyverse)
library(MASS)
library(ordinal)
```


Read in data: 
```{r}
data <- read_csv("ISSP_OECD2015.csv")
data

glimpse(data)
```

Drop rows with missing values, except for ISCO column. (Per PI)

```{r}
data_clean <- data%>%
  dplyr::select(-isco)%>%
  tidyr::drop_na()

data_clean$jobsec <- factor(data_clean$jobsec, ordered = TRUE, 
                            levels = c("Strongly disagree", "Disagree", "Neither nor", "Agree", "Strongly agree"))


#data_clean$relmgmt <- factor(data_clean$relmgmt)

data_clean$rmgt3 <- data_clean$relmgmt
data_clean$rmgt3[data_clean$relmgmt %in% c("Very bad", "Quite bad")] <- "Bad"
data_clean$rmgt3[data_clean$relmgmt %in% c("Very good", "Quite good")] <- "Good"
data_clean$rmgt3 <- factor(data_clean$rmgt3)

glimpse(data_clean)
```
Response, y = jobsec.  5 level ordinal response. 

Fixed effect with clm(): 

```{r}

fit0 <- clm(jobsec~ relmgmt + unemp +ltu + sex + age + sector + fulltime + union, data = data_clean)

fit1 <- clm(jobsec~ eprc + relmgmt + unemp +ltu + sex + age + sector + fulltime + union, data = data_clean)

fit2 <- clm(jobsec~ eprc * relmgmt + unemp +ltu + sex + age + sector + fulltime + union, data = data_clean)

summary(fit1)
summary(fit2)

anova(fit0, fit1, fit2)

anova(fit0, fit2)

```

Ordinal logistic regression mixed effects model with country as random effect, with clmm(): 

```{r}
fit3 <- clmm(jobsec~ eprc * relmgmt + unemp +ltu + sex + age + sector + fulltime + union + (1|country), data = data_clean)

summary(fit3)


fit4 <- clmm(jobsec~ eprc * rmgt3 + unemp +ltu + sex + age + sector + fulltime + union + (1|country), data = data_clean)

summary(fit4)
```


```{r}
AIC(fit1,fit2,fit3,fit4)
BIC(fit1,fit2,fit3,fit4)
```





Odds: 

bad neutral good, some way to combine odds? 

Job security 5 levels.  


Graphs.  3 graphs, 5 lines each.  Xaxis is eprc, y axis is probability of a specific level of job security. each graph corresponds to one level of relmgmgt. 

```{r}
library(effects)
Effect("eprc", fit4) %>% plot()



```


odds ratio plot proc logistic


```{r}
coef_table <- broom.mixed::tidy(fit4, conf.int = TRUE, conf.method = "Wald")


coef_table


coef_table <- coef_table %>%
  filter(!grepl("\\|", term)) %>%
  mutate(
    OR = exp(estimate),
    OR_low = exp(estimate - 1.96 * std.error),
    OR_high = exp(estimate + 1.96 * std.error),
    term = gsub(":", " × ", term)  
  )

coef_table


```


```{r}
library(ggplot2)

ggplot(coef_table, aes(x = OR, y = reorder(term, OR))) +
  geom_point(size = 3, color = "blue") +
  geom_errorbarh(aes(xmin = OR_low, xmax = OR_high), height = 0.2, color = "red") +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10() +
  labs(
    title = "Odds Ratios from Ordinal Logistic Regression (clmm)",
    x = "Odds Ratio (log scale)",
    y = ""
  ) +
  theme_minimal(base_size = 12)
```




```{r}
library(ggplot2)
library(gridExtra)



ggplot(coef_table, aes(y = term, x = OR)) +
  geom_point(shape = 18, size = 2) +  
  geom_errorbarh(aes(xmin = OR_low, xmax = OR_high), height = 0.25) +
  geom_vline(xintercept = 1, color = "red", linetype = "dashed", cex = 1, alpha = 0.5) +
  xlab("Odds Ratio (95% CI)") + 
  ylab(" ") + 
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x.bottom = element_text(size = 12, colour = "black"),
        axis.title.x = element_text(size = 12, colour = "black"))

coef_table

```







