---
title: "Analysis"
author: "Susie and Teddie"
format: pdf
---

Packages:

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(MASS)
library(ordinal)
```

Read in data:

```{r}
data <- read_csv("ISSP_OECD2015.csv")

glimpse(data)
```

Drop rows with missing values, and remove ISCO column. Collapse relationship with management into 3 levels: bad- comprised of Very Bad and Quite Bad, Neither Nor- comprised of just Neither Nor, good- comprised of Very Good and Quite Good. (Per PI)

```{r}
data_clean <- data%>%
  dplyr::select(-isco)%>%
  tidyr::drop_na()

data_clean$jobsec <- factor(data_clean$jobsec, ordered = TRUE, 
                            levels = c("Strongly disagree", "Disagree", "Neither nor", "Agree", "Strongly agree"))

data_clean$rmgt3 <- data_clean$relmgmt
data_clean$rmgt3[data_clean$relmgmt %in% c("Very bad", "Quite bad")] <- "Bad"
data_clean$rmgt3[data_clean$relmgmt %in% c("Very good", "Quite good")] <- "Good"
data_clean$rmgt3 <- factor(data_clean$rmgt3)

glimpse(data_clean)
```

Response, y = jobsec. 5 level ordinal response.

Fixed effect with clm():

```{r}

fit0 <- clm(jobsec~ relmgmt + unemp +ltu + sex + age + sector + fulltime + union, data = data_clean)

fit1 <- clm(jobsec~ eprc + relmgmt + unemp +ltu + sex + age + sector + fulltime + union, data = data_clean)

fit2 <- clm(jobsec~ eprc * relmgmt + unemp +ltu + sex + age + sector + fulltime + union, data = data_clean)

summary(fit1)
summary(fit2)

anova(fit0, fit1, fit2)

anova(fit0, fit2)

```

Ordinal logistic regression mixed effects model with country as random effect, with clmm():

```{r}
fit3 <- clmm(jobsec~ eprc * relmgmt + unemp +ltu + sex + age + sector + fulltime + union + (1|country), data = data_clean)

summary(fit3)


fit4 <- clmm(jobsec~ eprc * rmgt3 + unemp +ltu + sex + age + sector + fulltime + union + (1|country), data = data_clean)

summary(fit4)
```

```{r}
AIC(fit1,fit2,fit3,fit4)
BIC(fit1,fit2,fit3,fit4)
```

Odds:

bad neutral good, some way to combine odds?

Job security 5 levels.

Graphs. 3 graphs, 5 lines each. Xaxis is eprc, y axis is probability of a specific level of job security. each graph corresponds to one level of relmgmgt.

```{r}
library(effects)
Effect("eprc", fit4) %>% plot()

```

odds ratio plot proc logistic

```{r}
coef_table <- broom.mixed::tidy(fit4, conf.int = TRUE, conf.method = "Wald")


coef_table


coef_table <- coef_table %>%
  filter(!grepl("\\|", term)) %>%
  mutate(
    OR = exp(estimate),
    OR_low = exp(estimate - 1.96 * std.error),
    OR_high = exp(estimate + 1.96 * std.error),
    term = gsub(":", " × ", term)  
  )

coef_table


```

These estimates above are the log odds of each variable to the response. The 4 response ones are intercepts.

Plot logs odds, odds, or probabilities as a function of eprc for each level of rel mgmgt

to see how the odds change as eprc changes. This pattern is different or each level of relmgmt bc there is interaction.

```{r}
coef_table
```

$$
P(y_i \leq j) = \frac{exp(\theta_j - x_{i1}\beta_1 - \ldots - x_{ip}\beta_p)}{1 + exp(\theta_j - x_{i1}\beta_1 - \ldots - x_{ip}\beta_p)}
$$ 

$$
P(y_i = j) = P(y_i \leq j) - P(y_i \leq j-1), \hbox{ for } j=2,3,4,5
$$

```{r}
X <- model.matrix(jobsec~ eprc * rmgt3 + unemp +ltu + sex + age + sector + fulltime + union, data = data_clean)
head(X)
X %*% coef(fit4)
```

first create a new dataframe with the covariates I have in mind. fix continuous things to the mean of them, then keep eprc of min to max of data, then set rgmt to be a level, set demos to a fixed level. prediction is given what we fixed.

X\* the coeffs of the fit is going to give me the linear predictors. Then add the theta js for the different logits, then do formula for probability.

```{r}
names(data_clean)

library(VGAM)
fulltime2 <- ifelse(data_clean$fulltime=="Part-time", 1, 0)


mod<- vglm(jobsec~ eprc * rmgt3 + unemp +ltu + sex + age + sector + fulltime2 + union, data = data_clean, family = cumulative(ynames = TRUE, parallel = TRUE))


summary(mod)


```

```{r}
new <- data.frame(eprc = 0:6, rmgt3="Good", unemp = median(data_clean$unemp), 
                  ltu = median(data_clean$ltu), 
                  sex = "Male",
                  age = median(data_clean$age), sector = "Public sector", fulltime2 =  mean(fulltime2),
                  union = "Union member")
predict(mod, type = "response", newdata = new) %>% matplot(type="b")
```

```{r}
pr <- predict(mod, type = "response", newdata = new)
pr
str(pr)
```

One table with min and max values from where the probabilities start at eprc and go to at 6 eprc. 16 columns with 3 in each cell because. Thats what interaction means, for different levels of rgmt eprc is going to react differently.
